---
title: "Portfolio Practice"
author: "Team 8"
date: "4/12/2020"
output: pdf_document
---
# Creating A Portfolio From Machine Learning Algorithms
## Preparing The Data
```{r message=FALSE, warning=FALSE}
set.seed(123456)
library(tidyverse)
library(scorer)
library(glmnet)
library(gbm)
library(rpart)
library(randomForest)
library(janitor)
library(scales)


data <- read_csv("portfolio_data.csv")
data <- data[-c(1)]
data <- data %>% select(X1_1, everything())
# Deleting any prices from around the Corona Virus
data <- data[-c(200:505)]
data <- data[-c(2:130)]
data[2:ncol(data)] <- scale(data[2:ncol(data)])

## Split data for portfolio and for selecting model dropping ticker symbol and the tag 
smp_size <- floor(0.5 * nrow(data))
train_ind <- sample(seq_len(nrow(data)), size = smp_size)
model_data <- data[train_ind, ]
model_data <- model_data[-c(1)]
port_data <- data[-train_ind, ]


## Splt into Train Test
smp_size <- floor(0.75 * nrow(model_data))
train_ind <- sample(seq_len(nrow(model_data)), size = smp_size)
train <- model_data[train_ind, ]
test <- model_data[-train_ind, ]
test_real <- as.vector(test[[c(1)]])
```



## Finding The Best Algorithm
```{r}

results <- tibble(x = 0, model = 0, mse = 0)


for (x in 2:ncol(train)) {
  y <- train[c(1:x)]
  y_test <- test[c(2:x)]
  
  ## Linear
  fit <- lm(y$`10/2/2019` ~ ., data = y)
  fit_predict <- predict(fit, y_test)
  mse <- mean_squared_error(test_real, fit_predict)
  results <- results %>% add_row(x = x, model = "linear", mse = mse)
  
  ## Decision Tree
  tree <- rpart(y$`10/2/2019` ~ ., y)
  fit_predict <- predict(tree, y_test)
  mse <- mean_squared_error(test_real, fit_predict)
  results <- results %>% add_row(x = x, model = "Decision Tree", mse = mse)
  
  ## Random Forest
  y <- clean_names(y)
  y_test <- clean_names(y_test)
  rf <- randomForest(x10_2_2019 ~ ., y, ntree = 250, do.trace = F)
  fit_predict <- predict(rf, y_test)
  mse <- mean_squared_error(test_real, fit_predict)
  results <- results %>% add_row(x = x, model = "RForest", mse = mse)
 
  # Boosting
  boost <- gbm(x10_2_2019 ~ . , 
               data = y,
               n.trees = 250,
               shrinkage = .001,
               distribution = "gaussian")
  fit_predict <- predict.gbm(boost, y_test, 250)
  mse <- mean_squared_error(test_real, fit_predict)
  results <- results %>% add_row(x = x, model = "Boost", mse = mse)
}

for (x in 3:ncol(train)){
  y_matrix <- as.matrix(train[c(2:x)])
  y_test <- as.matrix(test[c(2:x)])
  y_train_real <- as.matrix(train[c(1)])
  
  lasso <- cv.glmnet(y_matrix, y_train_real, alpha = 1, nfolds = 10)
  fit_predict <- predict(lasso, y_test)
  mse <- mean_squared_error(test_real, fit_predict)
  results <- results %>% add_row(x = x, model = "Lasso", mse = mse)
}

for (x in 3:ncol(train)){
  y_matrix <- as.matrix(train[c(2:x)])
  y_test <- as.matrix(test[c(2:x)])
  y_train_real <- as.matrix(train[c(1)])
  
  ridge <- cv.glmnet(y_matrix, y_train_real, alpha = 0, nfolds = 10)
  fit_predict <- predict(ridge, y_test)
  mse <- mean_squared_error(test_real, fit_predict)
  results <- results %>% add_row(x = x, model = "Ridge", mse = mse)
}

results <- results[order(results$mse, decreasing = F),] 
results[2,]
View(results)
```
The results above show x, the best window size, model, the type of model used, and the MSE that was the best. 


## Final Model & Creating Predictions
```{r}
## Creating the Final Model
#### The Best performing model from above was simple linear regression with a window of 54
final_model_data <- model_data[1:55]
colnames(final_model_data) <- c(1:55)
final_model <- lm(final_model_data$`1` ~ ., data = final_model_data)

## Preparing The Portfolio Data
port_data <- port_data[1:56]
tickers <- port_data[c(1)]
port_data <- port_data[-c(1)]
colnames(port_data) <- c(1:55)
port_data <- cbind(tickers, port_data)

## Predicting For Portfolio Data
port_predictions <- predict(final_model, port_data)
port_data <- cbind(port_predictions, port_data)
port_data <- port_data %>%  mutate(spread = (port_predictions - port_data$`2`))
port_data <- port_data[order(port_data$spread, decreasing = T),] 
port_data <- port_data %>% mutate(actual_spread = (port_data$`1` - port_data$`2`))
```



## The Winning Basket
For this exercise we are just picking the top predictions that would gain us the most money. Meaning The difference between the predicition and the day before that would give use the most money in the aggregate. We are not looking at the price of the stock, but rather an entity that would gain the most money in the aggregate. 
```{r}
### Winning Basket
winning_basket <- port_data[1:5,]
x <- winning_basket[,1:3]
colnames(x) <- c("Predictions", "Tickers", "Actual Price")
x <- x %>% select(Tickers, Predictions, `Actual Price`)
x
```

## Creating Random Baskets
These baskets are randomly created to compete against our predicted winning Basket. 
```{r}
### Random Baskets For Comparision 
rb1 <- port_data %>% sample_n(5)
set.seed(2)
rb2 <- port_data %>% sample_n(5)
set.seed(3)
rb3 <- port_data %>% sample_n(5)
set.seed(4)
rb4 <- port_data %>% sample_n(5)
set.seed(5)
rb5 <- port_data %>% sample_n(5)
set.seed(6)
rb6 <- port_data %>% sample_n(5)
set.seed(7)
rb7 <- port_data %>% sample_n(5)
set.seed(8)
rb8 <- port_data %>% sample_n(5)
set.seed(9)
rb9 <- port_data %>% sample_n(5)
set.seed(10)
rb10 <- port_data %>% sample_n(5)
```
 
## Creating A Dataframe & Data
```{r}
### Gains/Losses
Gains_Losses <- c(sum(winning_basket$actual_spread),
                   sum(rb1$actual_spread),
                   sum(rb2$actual_spread),
                   sum(rb3$actual_spread),
                   sum(rb4$actual_spread),
                   sum(rb5$actual_spread),
                   sum(rb6$actual_spread),
                   sum(rb7$actual_spread),
                   sum(rb8$actual_spread),
                   sum(rb9$actual_spread),
                   sum(rb10$actual_spread))

### Average Gains/Losses
Average <- c(sum(winning_basket$actual_spread)/5,
                  sum(rb1$actual_spread)/5,
                  sum(rb2$actual_spread)/5,
                  sum(rb3$actual_spread)/5,
                  sum(rb4$actual_spread)/5,
                  sum(rb5$actual_spread)/5,
                  sum(rb6$actual_spread)/5,
                  sum(rb7$actual_spread)/5,
                  sum(rb8$actual_spread)/5,
                  sum(rb9$actual_spread)/5,
                  sum(rb10$actual_spread)/5)

### Portfolio Names
Name <- c('Winner',
               'Basket1',
               'Basket2',
               'Basket3',
               'Basket4',
               'Basket5',
               'Basket6',
               'Basket7',
               'Basket8',
               'Basket9',
               'Basket10')

Portfolio <- data.frame(Name, Gains_Losses, Average)
Portfolio <- Portfolio %>%  mutate(Beat_Winner = if_else(Gains_Losses > Portfolio[1,2], "Yes", "No"))
Portfolio <- Portfolio %>%  
  mutate(Beat_Aggregate = if_else(Average > mean(port_data$actual_spread), "Yes", "No"))

Portfolio
```


# An Analysis
## Winning Basket 
```{r}
Portfolio[1,]
```
The above output shows the winning basket's proporties. The real Gains_Losses shows the loss incurred for the portfolio. The average shows the average loss of the portfolio for all the stocks. The Beat Winner variables is useless as it is the winner. The Beat Aggregate shows that it did not beat the Aggregate, the aggregate is just the average of all of the prices difference between the last day and the current day. It was then compared against the average of the portfolios returns. 


## Random Baskets
Removing the Winning Basket and converting data back into numerical data for some statistics. 
```{r}
### Doing Some Analysis
Analysis <- Portfolio[2:11,]
Analysis <- Analysis %>%  mutate(Beat_Winner = if_else(Beat_Winner == "Yes", 1, 0))
Analysis <- Analysis %>%  mutate(Beat_Aggregate = if_else(Beat_Aggregate == "Yes", 1, 0))
```


### Amount That Beat The Winning Portfolio
70 percent of the randomly put together portfolios are beating our winning portfolio.   
```{r}
percent(mean(Analysis$Beat_Winner))
```

### Amount That Beat The Aggregate
40 percent of the randomly put together portfolios are beating the aggregate.
```{r}
percent(mean(Analysis$Beat_Aggregate))
```

